Modeling workflow: Lake $N_2O$ survey data

### 2023-03-08 scripts/NLA_N2O_models.Rmd
Revisions to rmd text and format. Adding github_document/.md version

---
title: "Bayesian models for NLA 2017 $N_2O$ survey data"
author: "Roy Martin, Jake Beaulieu, Michael McManus"
date: "`r Sys.Date()`"
output:
  html_notebook:
    toc: yes
    toc_depth: 4
    toc_float: yes
    code_folding: show
    font_size: 14
    number_sections: yes
    theme: simplex
bibliography: RWM_Endnote_Library.bib
link-citations: yes
editor_options:
  chunk_output_type: inline
---

@@

# Rationale and Objectives
This document details the modeling workflow implemented for estimating dissolved and equilibrium N2O concentrations and saturation ratios using the 2017 Nation Lakes Assessment (NLA) survey data. The NLA sampling sites were distributed among the target population of US lakes (in the lower 48 states) according to a probabilistic survey design with samples stratified among categories of lake surface area, WSA9 ecoregion, and US state (excluding AK and HI). Due to the stratification scheme, some types of lakes in the sample population were intentionally over-represented (e.g., large lakes) and some were under-represented (e.g., small lakes) relative to the target population. Due to the unequal probability design, inferences from the sample had to be adjusted for inferences on the broader populations of interest (e.g, National-, state-, ecoregion-, and size class-specific estimates).

The concept of the "complete data likelihood" is useful for conceptualizing biases arising from sampling design [@Zachmann_etal_2022; @Gelman_etal_2014 Ch. 8; @Link_Barker_2010]. For the NLA survey data, the population of US lakes in the lower 48 states larger than 4 hectares was considered the complete data and the probabilistic samples were considered a subset of that complete data. The portion of US lakes not included in the sample were considered "missing" from the complete data _not_ at random, but conditional on the pre-specified design (stratification) variables. This non-random missingness was not ignorable for the purpose of making inferences from the sample to the target population. In a model-based framework, however, including the design parameters as predictors in a regression model is one way to adjust for the missingness. For a thorough and recent treatment of this concept in the context of national surveys of environmental resources, refer to [@Zachmann_etal_2022]. This concept is a key motivator for the increasingly popular mulitilevel regression with poststratification (MRP) approach to model-based inference [@Gelman_etal_2014; Gelmant_etal_2020 Ch. 17].

The following workflow illustrates our model-based approach, based largely on the logic of MRP, but with an elaboration on the poststratification step to enable eventual estimates of total gas flux at the population level, which required scaling up from lake-level estimates. The typical MRP process is carried out in two steps. The first step is to fit regression models for the response variables of interest (e.g., dissolved N2O, equilibrium N2O) conditional on the survey design variables ┼(i.e., ecoregion, state, lake size). The second step is post-stratfication, wherein the posterior parameter estimates from the regression model for the sample population are weighted based on their known or assumed distribution in the population of interest [i.e., post-stratification table; @Gelman_etal_2020 Ch. 17]. The poststratification table in our case, for example, would be a population summary of lakes among the design variables: ecoregion, state, and size category. However, because we eventually needed lake-level estmates, instead of predicting to a postratification table, we predicted to each individual lake in the population of interest. This meant predicting to the full target population of 465,897 natural and man made US lakes larger than 4 hectares in the lower 48 states. These predictions were assumed relevant to average conditions during the "index period" for each lake in 2017. Details about the sampling frame as well as the target population are further clarified in the workbook below with data summaries and code.

For the regressions, we used multilevel models fit in a fully Bayesian fashion Multilevel models are thought to work well in this context because they provide regularized estimates along the design groupings, which can improve out-of-sample inferences [@McElreath_2020]. Inferences for lake types that may be missing from the sample, but are part of the population of interest are also straightforward using this approach [@Gelman_etal_2020 Ch. 17; @McElreath_2020]. More information these models, their specific parameters, R code, fit evaluations, and resulting inferences are presented in this document.

The overriding objective of the modeling effort was to provide population level estimates for (1) dissolved and equilibrium N2O concentrations; (2) the N2O saturation ratio (i.e., dissolved N2O/equilibrium N2O); and (3) the proportion of under-saturated water bodies (i.e., saturation ratio < 1). The estimates would also be used to later estimate the total flux of N2O gas attributable to the target population of lakes over the index period. The saturation ratio estimates were calculated as a derived quantity based on the ratio of modeled dissolved to equilibrium N2O. Because dissolved and equilibrium N2O were observed on the same sample units (lake sites), we developed models for estimating their joint distribution. The response variable in the models was, therefore, multivariate to account for potential statistical dependencies between dissolved and equilibrium N2O due to, for example, common dependencies on geography. Although point predictions of the mean marginal probabilities from separate models could be comparable, a joint model allowing correlated observation-level errors (i.e., residuals) was expected to better capture uncertainty and potentially improve out-of-sample predictions, should the variables be conditionally correlated [@Warton_etal_2015; @Poggiato_etal_2021]. All of the models fit were constructed using the `brms` package [@Burkner_2017] in `R` [@R_Core_Team_2021] as an interface to Stan, a software package for fitting fully Bayesian models via Hamiltonian Monte Carlo [HMC; @Stan_Development_Team_2018_a; @Stan_Development_Team_2018_b; @Stan_Development_Team_2018_c].

@@

print(df_model)
```

@@

print(df_test)
```

@@

print(sframe)
```

@@

print(pframe)
```

@@

print(samp_props)
```

@@

$$ N_2O_{dissolved}\sim 1 + (1|a|WSA9) + (1|b|WSA9:state) + (1|c|WSA9:state:size) \\ 
N_2O_{equilibrium}\sim 1 + (1|a|WSA9) + (1|b|WSA9:state) + (1|c|WSA9:state:size)$$

@@

Before investing too much into the interpretation of this model, however, the model fit was evaluated below using a series of graphical posterior predictive checks [PPC; @Gelman_etal_2014; @Gelman_etal_2020, Ch. 11].

### Model checks
#### Dissolved N2O
Below are a series of panels illustrating graphical PPCs for the log(N2O) component of the model. The top left panel compares a density plot of the observed data (black line) to density lines drawn for 200 samples from the posterior predictive distribution (PPD; blue lines) of the fitted model. The top right panel similarly compares the cumulative density distributions. The left middle panel simulataneously compares means _vs._ standard deviations for 1000 draws from the PPD (blue dots) to the sample mean and standard deviation (black dot). The right middle panel compares skewness _vs._ kurtosis for 1000 draws from the PPD to the skewness and kurtosis values calculated for the observed data. The bottom left panel compares max _vs._ min values for 1000 draws from the PPD to the max and min values of the sample data. Finally, the bottom right panel shows the observed _vs._ average predicted values for each observation in the sample. The average predicted values were calculated as the mean prediction for each observation in the PPD based on 1000 draws.
```{r ppc_n2o1, echo=FALSE, fig.align='center', fig.height=8, fig.width=8, message=FALSE, warning=FALSE}
load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/n2o_mod1.rda")
grid.arrange(
pp_check(n2o_mod1, 
         resp = "logn2o",
         type = "dens_overlay",
         ndraws = 200,
         cores = 1) + 
  theme_tidybayes() +
  xlab(expression(paste("log(N"[2],"O) concentration"))) + 
  ylab("density")
,
pp_check(n2o_mod1, 
         resp = "logn2o",
         type = "ecdf_overlay",
         ndraws = 200, 
         cores = 1) + 
  theme_tidybayes() +
  xlab(expression(paste("log(N"[2],"O) concentration"))) + 
  ylab("cumulative density")
,
pp_check(n2o_mod1, 
         resp = "logn2o",
         type = "stat_2d", 
         stat = c("mean", "sd"),
         ndraws = 1000,
         cores = 1) + 
  theme_tidybayes()
,
pp_check(n2o_mod1, 
         resp = "logn2o",
         type = "stat_2d", 
         stat = c("kurtosis", "skewness"), 
         ndraws = 1000,
         cores = 1) + 
  theme_tidybayes()
,
pp_check(n2o_mod1, 
         resp = "logn2o",
         type = "stat_2d", 
         stat = c("min", "max"), 
         ndraws = 1000,
         cores = 1) + 
  theme_tidybayes()
,
pp_check(n2o_mod1, 
         resp = "logn2o",
         type = "scatter_avg",
         ndraws = 1000,
         cores = 1) + 
  theme_tidybayes()
, ncol = 2)
```
The general takeaway from the PPCs above was that the model replicated the central tendency of the observed data fairly well, but failed to sufficiently replicate other important aspects of the distribution, such as skewness and kurtosis. The observed _vs._ average predictions scatterplot suggested substantial heteroscedasticity in the errors.

The same checks were run below, but for the test set of 95 held-out, second-visit data points. 
```{r ppc_n2o1_test, echo=FALSE, fig.align='center', fig.height=8, fig.width=8, message=FALSE, warning=FALSE}
load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/n2o_mod1.rda")
load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/df_test.rda")
grid.arrange(
pp_check(n2o_mod1, 
         newdata = df_test,
         resp = "logn2o",
         type = "dens_overlay",
         ndraws = 200,
         cores = 1) + 
  theme_tidybayes() +
  xlab(expression(paste("log(N"[2],"O) concentration"))) + 
  ylab("density")
,
pp_check(n2o_mod1, 
         newdata = df_test,
         resp = "logn2o",
         type = "ecdf_overlay",
         ndraws = 200, 
         cores = 1) + 
  theme_tidybayes() +
  xlab(expression(paste("log(N"[2],"O) concentration"))) + 
  ylab("cumulative density")
,
pp_check(n2o_mod1, 
         newdata = df_test,
         resp = "logn2o",
         type = "stat_2d", 
         stat = c("mean", "sd"),
         ndraws = 1000,
         cores = 1) + 
  theme_tidybayes()
,
pp_check(n2o_mod1, 
         newdata = df_test,
         resp = "logn2o",
         type = "stat_2d", 
         stat = c("kurtosis", "skewness"), 
         ndraws = 1000,
         cores = 1) + 
  theme_tidybayes()
,
pp_check(n2o_mod1, 
         newdata = df_test,
         resp = "logn2o",
         type = "stat_2d", 
         stat = c("min", "max"), 
         ndraws = 1000,
         cores = 1) + 
  theme_tidybayes()
,
pp_check(n2o_mod1, 
         newdata = df_test,
         resp = "logn2o",
         type = "scatter_avg",
         ndraws = 1000,
         cores = 1) + 
  theme_tidybayes()
, ncol = 2)
```
The patterns in misfit indicated above for the re-visit data were similar to the patterns indicated in the PPCs with the training data.

#### Equilibrium N2O
Below are PPCs for the equilibrium N2O component of the model. As with the dissolved N2O response above, the model did an OK job at replicating the central tendency, but performed less well at replicating some important aspects of the overall distribution. 
```{r ppc_n2oeq1, echo=FALSE, fig.align='center', fig.height=8, fig.width=8, message=FALSE, warning=FALSE}
load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/n2o_mod1.rda")
grid.arrange(
pp_check(n2o_mod1, 
         resp = "logn2oeq",
         type = "dens_overlay",
         ndraws = 200,
         cores = 1) + 
  theme_tidybayes() +
  xlab(expression(paste("log(N"[2],"O) concentration"))) + 
  ylab("density")
,
pp_check(n2o_mod1, 
         resp = "logn2oeq",
         type = "ecdf_overlay",
         ndraws = 200, 
         cores = 1) + 
  theme_tidybayes() +
  xlab(expression(paste("log(N"[2],"O) concentration"))) + 
  ylab("cumulative density")
,
pp_check(n2o_mod1, 
         resp = "logn2oeq",
         type = "stat_2d", 
         stat = c("mean", "sd"),
         ndraws = 1000,
         cores = 1) + 
  theme_tidybayes()
,
pp_check(n2o_mod1, 
         resp = "logn2oeq",
         type = "stat_2d", 
         stat = c("kurtosis", "skewness"),
         ndraws = 1000,
         cores = 1) + 
  theme_tidybayes()
,
pp_check(n2o_mod1, 
         resp = "logn2oeq",
         type = "stat_2d", 
         stat = c("min", "max"), 
         ndraws = 1000,
         cores = 1) + 
  theme_tidybayes()
,
pp_check(n2o_mod1, 
         resp = "logn2oeq",
         type = "scatter_avg", 
         ndraws = 1000,
         cores = 1) + 
  theme_tidybayes()
, ncol = 2)
```

Below are the same PPCs for equilibrium N2O in the re-visit sites.
```{r ppc_n2oeq1_test, echo=FALSE, fig.align='center', fig.height=8, fig.width=8, message=FALSE, warning=FALSE}
load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/n2o_mod1.rda")
grid.arrange(
pp_check(n2o_mod1,
         newdata = df_test,
         resp = "logn2oeq",
         type = "dens_overlay",
         ndraws = 200,
         cores = 1) + 
  theme_tidybayes() +
  xlab(expression(paste("log(N"[2],"O) concentration"))) + 
  ylab("density")
,
pp_check(n2o_mod1, 
         newdata = df_test,
         resp = "logn2oeq",
         type = "ecdf_overlay",
         ndraws = 200, 
         cores = 1) + 
  theme_tidybayes() +
  xlab(expression(paste("log(N"[2],"O) concentration"))) + 
  ylab("cumulative density")
,
pp_check(n2o_mod1, 
         newdata = df_test,
         resp = "logn2oeq",
         type = "stat_2d", 
         stat = c("mean", "sd"),
         ndraws = 1000,
         cores = 1) + 
  theme_tidybayes()
,
pp_check(n2o_mod1, 
         newdata = df_test,
         resp = "logn2oeq",
         type = "stat_2d", 
         stat = c("kurtosis", "skewness"),
         ndraws = 1000,
         cores = 1) + 
  theme_tidybayes()
,
pp_check(n2o_mod1, 
         newdata = df_test,
         resp = "logn2oeq",
         type = "stat_2d", 
         stat = c("min", "max"), 
         ndraws = 1000,
         cores = 1) + 
  theme_tidybayes()
,
pp_check(n2o_mod1, 
         newdata = df_test,
         resp = "logn2oeq",
         type = "scatter_avg", 
         ndraws = 1000,
         cores = 1) + 
  theme_tidybayes()
, ncol = 2)
```
#### Bivariate
The graphical check below compares bivariate density contours estimated from the observed data (black lines) to density contours estimated for each of 20 draws from the PPD. The model appeared to do a good job of replicating the bivariate mean, but was poor at representing the overall joint distribution.
```{r ppc_biv1, echo=FALSE, fig.align='center', fig.height=4, fig.width=4, message=FALSE, warning=FALSE}
load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/n2o_mod1.rda")
df_model %>%
  add_predicted_draws(n2o_mod1, 
                      ndraws = 20) %>%
  tidyr::pivot_wider(names_from = .category,
              values_from = .prediction) %>%
  ggplot(aes(x = log(n2o), y = log(n2o_eq))) +
  geom_density_2d(aes(x = logn2o, 
                      y = logn2oeq, 
                      group = .draw),
                  bins = 10,
                  color = "lightblue", 
                  alpha = 0.4) +
  geom_density_2d(color = "black", bins = 10) +
  xlim(1.25, 2.75) +
  ylim(1.25, 2.75) +
  geom_abline(intercept = 0, slope = 1, linetype = "dashed") +
  theme_tidybayes()
```

The same bivariate check is shown below for the re-visit data.
```{r ppc_biv1_test, echo=FALSE, fig.align='center', fig.height=4, fig.width=4, message=FALSE, warning=FALSE}
load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/n2o_mod1.rda")
df_test %>%
  add_predicted_draws(n2o_mod1, 
                      ndraws = 20) %>%
  tidyr::pivot_wider(names_from = .category,
              values_from = .prediction) %>%
  ggplot(aes(x = log(n2o), y = log(n2o_eq))) +
  geom_density_2d(aes(x = logn2o, 
                      y = logn2oeq, 
                      group = .draw),
                  bins = 10,
                  color = "lightblue", 
                  alpha = 0.4) +
  geom_density_2d(color = "black", bins = 10) +
  xlim(1.25, 2.75) +
  ylim(1.25, 2.75) +
  geom_abline(intercept = 0, slope = 1, linetype = "dashed") +
  theme_tidybayes()
```

@@

$$N_2O_{dissolved}:N_2O_{equilibrium}$$

@@

```{r ppc_sat1, echo=FALSE, fig.align='center', fig.height=8, fig.width=8, message=FALSE, warning=FALSE}

@@

```{r ppc_sat2, echo=FALSE, fig.align='center', fig.height=8, fig.width=8, message=FALSE, warning=FALSE}

@@

```{r ppc_sat_check_mod_n2o3, echo=FALSE, fig.align='center', fig.height=8, fig.width=8, message=FALSE, warning=FALSE}

@@

```{r ppc_sat_check_mod_n2o6, echo=FALSE, fig.align='center', fig.height=8, fig.width=8, message=FALSE, warning=FALSE}

>>>>

---
title: "Modeling workflow: Lake $N_2O$ survey data"
author: "Roy Martin, Jake Beaulieu, Michael McManus"
date: "`r Sys.Date()`"
output:
  github_document:
    toc: TRUE
    toc_depth: 4
    number_sections: TRUE
    df_print: "tibble"
    math_method: 
      engine: webtex
      url: https://latex.codecogs.com/svg.image?
    html_preview: TRUE
    keep_html: TRUE
bibliography: RWM_Endnote_Library.bib
link-citations: yes
editor_options:
  chunk_output_type: inline
---

@@

# Background and Objectives
This document details the modeling workflow implemented for estimating dissolved and equilibrium N2O gas concentrations and saturation ratios for a targeted population of US lakes, which included all freshwater lakes and reservoirs in the lower 48 US states larger than 4 ha. Data for these estimates came from the 2017 Nation Lakes Assessment (NLA) survey, wherein waterbodies were sampled according to a stratified, unequal probability design. Sampling was random within categories of lake size (surface area in hectares), WSA9 ecoregion, and US state (excluding AK and HI). In addition, larger lakes were intentionally over-sampled in proportion to smaller lakes. In order to make inferences from this sample to the population of interest, the estimates needed to be adjusted according to these factors.

The "Complete Data Likelihood" is a useful conceptualization for describing how biases arise in sampling [@Link_Barker_2010; @Zachmann_etal_2022]. For the 2017 NLA survey, the complete data, or population of interest, was all US lakes and reservoirs in the lower 48 states larger than 4 hectares. In this context, the survey samples were considered a subset of the complete data. Due to the pattern of selection assigned by the survey, some lakes of certain categories (e.g., small lakes) in the target population were missing from the sample not at random, but conditional on the unequal probability design. This type of non-random "missingness" is not ignorable when making inferences from a sample to a population of interest, as it can lead to biased estimates [@Gelman_etal_2014, Ch. 8].

A relatively simple, model-based strategy to account for this type of selection bias is to include the survey design variables as predictors in a regression model and then poststratify the estimates based on a known target population distribution; an adjustment procedure often referred to as multilevel regression and poststratification [MRP, @Park_etal_2004; @Gelman_Little_1997; @Gelman_etal_2020, Ch. 17]. With a stratified design, such as the one used for the 2017 NLA, the selection scheme results in pre-specified groupings within which samples may be drawn at random. Within a particular grouping or stratification, the selected observations would be considered a simple random sample (SRS) from that group's population. A regression model conditioning the responses on that grouping structure would then result in group-specific estimates that are unbiased with respect to the groups. To make make unbiased inferences to a broader population of interest, those regression estimates may be re-weighted by the known distribution of units from those same groups in a broader target population. That distribution may be summarized in a poststratification table, which, for the 2017 NLA survey, could be a cross-tabulation of lakes in the target population along the design variables: ecoregion, state, and size category. Thus, a regression model conditioning on this design would provide estimates for groups of lakes described by a size classification within a particular state and ecoregion. To adjust the regression estimates to an estimate for the population of interest (e.g., all US lakes larger than 4 ha), those group-specific regression estimates from the sample would be re-weighted based on the proportional distribution of lakes in those groups in the target population. For a recent applied example and tutorial, see @Kennedy_Gelman_2021. For an example of a recent application in the context of national surveys of environmental resources, see @Zachmann_etal_2022.

To make population-level estimates from the 2017 NLA dissolved gas data, a model-based approach, similar to those referenced above, was followed. The specific workflow, data, models, and code used to make these estimates are documented in the remainder of this document. First, Bayesian multilevel regression models were used to fit the sample data to the design variable structure. Multilevel models are often recommended in this type of survey estimation because they provide regularized estimates along the design groupings, which can improve out-of-sample inferences [e.g., @Gelman_Little_1997; @Kennedy_Gelman_2021]. Inferences for group levels that may be missing from the sample, but are part of the population of interest are also straightforward using this approach [@Gelman_etal_2020 Ch. 17; @McElreath_2020]. The second step is typically a simple poststratification of the sample-based estimates over a table of target population weights. However, the estimates ultimately needed to be useful for estimating the total gas flux at the population level, which required lake level estimates. That is, flux estimates were dependent on lake-level measures of surface area. Therefore, instead of predicting to a postratification table, predictions were made to each individual lake in the population of interest. This meant predicting to 465,897 individual natural and man made US lakes larger than 4 hectares in the lower 48 states. These predictions were assumed relevant to average conditions during the biological index period for each lake in 2017.

The specific objectives of the modeling effort were to provide population level estimates for (1) dissolved and equilibrium N2O concentrations; (2) the N2O saturation ratio (i.e., dissolved N2O/equilibrium N2O); and (3) the proportion of under-saturated water bodies (i.e., saturation ratio < 1). The gas estimates would also be subsequently used to estimate the total flux of N2O gas attributable to the target population of lakes over the index period. The saturation ratio estimates were modeled as a derived quantity from the joint model of dissolved and equilibrium N2O and were calculated as the ratio of modeled dissolved to equilibrium N2O. Because dissolved and equilibrium N2O were observed on the same sample units (lake sites), models for estimating their joint distribution were used. That is, the response variable was multivariate to account for any potential statistical dependencies between dissolved and equilibrium N2O due to, for example, common dependencies on geography. Although point predictions of the mean marginal probabilities from separate models could provide comparable estimates, a joint model allowing correlated errors was expected to better capture uncertainty and potentially improve out-of-sample predictions, should the variables be conditionally correlated [@Warton_etal_2015; @Poggiato_etal_2021]. All of the models fit were constructed using the `brms` package [@Burkner_2017] in `R` [@R_Core_Team_2021] as an interface to Stan, a software package for fitting fully Bayesian models via Hamiltonian Monte Carlo [HMC, @Stan_Development_Team_2018_a; @Stan_Development_Team_2018_b; @Stan_Development_Team_2018_c]. More specific details on these models is provided in the "Model fitting" section below.

@@

df_model %>%
  print()
```

@@

df_test %>%
  print()
```

@@

sframe %>%
  print()
```

@@

pframe %>%
  print()
```

@@

samp_props %>%
  print()
```

@@

$$ N_2O_{diss}\sim 1 + (1|a|WSA9) + (1|b|WSA9:state) + (1|c|WSA9:state:size) \\ 
N_2O_{equi}\sim 1 + (1|a|WSA9) + (1|b|WSA9:state) + (1|c|WSA9:state:size)$$

@@

Before investing too much into the interpretation of this model, however, the model fit was evaluated below using a series of graphical posterior predictive checks [PPC, @Gelman_etal_2014; @Gelman_etal_2020, Ch. 11].

### Model checks
#### Dissolved N2O
Below are a series of panels illustrating graphical PPCs for the log(N2O) component of the model. The top left panel compares a density plot of the observed data (black line) to density lines drawn for 200 samples from the posterior predictive distribution (PPD; blue lines) of the fitted model. The top right panel similarly compares the cumulative density distributions. The left middle panel simulataneously compares means _vs._ standard deviations for 1000 draws from the PPD (blue dots) to the sample mean and standard deviation (black dot). The right middle panel compares skewness _vs._ kurtosis for 1000 draws from the PPD to the skewness and kurtosis values calculated for the observed data. The bottom left panel compares max _vs._ min values for 1000 draws from the PPD to the max and min values of the sample data. Finally, the bottom right panel shows the observed _vs._ average predicted values for each observation in the sample. The average predicted values were calculated as the mean prediction for each observation in the PPD based on 1000 draws.
```{r ppc_n2o1, echo=FALSE, fig.align='center', fig.height=8, fig.width=8, message=FALSE, warning=FALSE, cache=TRUE}
load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/n2o_mod1.rda")
grid.arrange(
pp_check(n2o_mod1, 
         resp = "logn2o",
         type = "dens_overlay",
         ndraws = 200,
         cores = 1) + 
  theme_tidybayes() +
  xlab(expression(paste("log(N"[2],"O) concentration"))) + 
  ylab("density")
,
pp_check(n2o_mod1, 
         resp = "logn2o",
         type = "ecdf_overlay",
         ndraws = 200, 
         cores = 1) + 
  theme_tidybayes() +
  xlab(expression(paste("log(N"[2],"O) concentration"))) + 
  ylab("cumulative density")
,
pp_check(n2o_mod1, 
         resp = "logn2o",
         type = "stat_2d", 
         stat = c("mean", "sd"),
         ndraws = 1000,
         cores = 1) + 
  theme_tidybayes()
,
pp_check(n2o_mod1, 
         resp = "logn2o",
         type = "stat_2d", 
         stat = c("kurtosis", "skewness"), 
         ndraws = 1000,
         cores = 1) + 
  theme_tidybayes()
,
pp_check(n2o_mod1, 
         resp = "logn2o",
         type = "stat_2d", 
         stat = c("min", "max"), 
         ndraws = 1000,
         cores = 1) + 
  theme_tidybayes()
,
pp_check(n2o_mod1, 
         resp = "logn2o",
         type = "scatter_avg",
         ndraws = 1000,
         cores = 1) + 
  theme_tidybayes()
, ncol = 2)
```
The general takeaway from the PPCs above was that the model replicated the central tendency of the observed data fairly well, but failed to sufficiently replicate other important aspects of the distribution, such as skewness and kurtosis. The observed _vs._ average predictions scatterplot suggested substantial heteroscedasticity in the errors.

The same checks were run below, but for the test set of 95 held-out, second-visit data points. 
```{r ppc_n2o1_test, echo=FALSE, fig.align='center', fig.height=8, fig.width=8, message=FALSE, warning=FALSE, cache=TRUE}
load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/n2o_mod1.rda")
load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/df_test.rda")
grid.arrange(
pp_check(n2o_mod1, 
         newdata = df_test,
         resp = "logn2o",
         type = "dens_overlay",
         ndraws = 200,
         cores = 1) + 
  theme_tidybayes() +
  xlab(expression(paste("log(N"[2],"O) concentration"))) + 
  ylab("density")
,
pp_check(n2o_mod1, 
         newdata = df_test,
         resp = "logn2o",
         type = "ecdf_overlay",
         ndraws = 200, 
         cores = 1) + 
  theme_tidybayes() +
  xlab(expression(paste("log(N"[2],"O) concentration"))) + 
  ylab("cumulative density")
,
pp_check(n2o_mod1, 
         newdata = df_test,
         resp = "logn2o",
         type = "stat_2d", 
         stat = c("mean", "sd"),
         ndraws = 1000,
         cores = 1) + 
  theme_tidybayes()
,
pp_check(n2o_mod1, 
         newdata = df_test,
         resp = "logn2o",
         type = "stat_2d", 
         stat = c("kurtosis", "skewness"), 
         ndraws = 1000,
         cores = 1) + 
  theme_tidybayes()
,
pp_check(n2o_mod1, 
         newdata = df_test,
         resp = "logn2o",
         type = "stat_2d", 
         stat = c("min", "max"), 
         ndraws = 1000,
         cores = 1) + 
  theme_tidybayes()
,
pp_check(n2o_mod1, 
         newdata = df_test,
         resp = "logn2o",
         type = "scatter_avg",
         ndraws = 1000,
         cores = 1) + 
  theme_tidybayes()
, ncol = 2)
```
The patterns in misfit indicated above for the re-visit data were similar to the patterns indicated in the PPCs with the training data.

#### Equilibrium N2O
Below are PPCs for the equilibrium N2O component of the model. As with the dissolved N2O response above, the model did an OK job at replicating the central tendency, but performed less well at replicating some important aspects of the overall distribution. 
```{r ppc_n2oeq1, echo=FALSE, fig.align='center', fig.height=8, fig.width=8, message=FALSE, warning=FALSE, cache=TRUE}
load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/n2o_mod1.rda")
grid.arrange(
pp_check(n2o_mod1, 
         resp = "logn2oeq",
         type = "dens_overlay",
         ndraws = 200,
         cores = 1) + 
  theme_tidybayes() +
  xlab(expression(paste("log(N"[2],"O) concentration"))) + 
  ylab("density")
,
pp_check(n2o_mod1, 
         resp = "logn2oeq",
         type = "ecdf_overlay",
         ndraws = 200, 
         cores = 1) + 
  theme_tidybayes() +
  xlab(expression(paste("log(N"[2],"O) concentration"))) + 
  ylab("cumulative density")
,
pp_check(n2o_mod1, 
         resp = "logn2oeq",
         type = "stat_2d", 
         stat = c("mean", "sd"),
         ndraws = 1000,
         cores = 1) + 
  theme_tidybayes()
,
pp_check(n2o_mod1, 
         resp = "logn2oeq",
         type = "stat_2d", 
         stat = c("kurtosis", "skewness"),
         ndraws = 1000,
         cores = 1) + 
  theme_tidybayes()
,
pp_check(n2o_mod1, 
         resp = "logn2oeq",
         type = "stat_2d", 
         stat = c("min", "max"), 
         ndraws = 1000,
         cores = 1) + 
  theme_tidybayes()
,
pp_check(n2o_mod1, 
         resp = "logn2oeq",
         type = "scatter_avg", 
         ndraws = 1000,
         cores = 1) + 
  theme_tidybayes()
, ncol = 2)
```

Below are the same PPCs for equilibrium N2O in the re-visit sites.
```{r ppc_n2oeq1_test, echo=FALSE, fig.align='center', fig.height=8, fig.width=8, message=FALSE, warning=FALSE, cache=TRUE}
load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/n2o_mod1.rda")
grid.arrange(
pp_check(n2o_mod1,
         newdata = df_test,
         resp = "logn2oeq",
         type = "dens_overlay",
         ndraws = 200,
         cores = 1) + 
  theme_tidybayes() +
  xlab(expression(paste("log(N"[2],"O) concentration"))) + 
  ylab("density")
,
pp_check(n2o_mod1, 
         newdata = df_test,
         resp = "logn2oeq",
         type = "ecdf_overlay",
         ndraws = 200, 
         cores = 1) + 
  theme_tidybayes() +
  xlab(expression(paste("log(N"[2],"O) concentration"))) + 
  ylab("cumulative density")
,
pp_check(n2o_mod1, 
         newdata = df_test,
         resp = "logn2oeq",
         type = "stat_2d", 
         stat = c("mean", "sd"),
         ndraws = 1000,
         cores = 1) + 
  theme_tidybayes()
,
pp_check(n2o_mod1, 
         newdata = df_test,
         resp = "logn2oeq",
         type = "stat_2d", 
         stat = c("kurtosis", "skewness"),
         ndraws = 1000,
         cores = 1) + 
  theme_tidybayes()
,
pp_check(n2o_mod1, 
         newdata = df_test,
         resp = "logn2oeq",
         type = "stat_2d", 
         stat = c("min", "max"), 
         ndraws = 1000,
         cores = 1) + 
  theme_tidybayes()
,
pp_check(n2o_mod1, 
         newdata = df_test,
         resp = "logn2oeq",
         type = "scatter_avg", 
         ndraws = 1000,
         cores = 1) + 
  theme_tidybayes()
, ncol = 2)
```

#### Bivariate
The graphical check below compares bivariate density contours estimated from the observed data (black lines) to density contours estimated for each of 20 draws from the PPD. The model appeared to do a good job of replicating the bivariate mean, but was poor at representing the overall joint distribution.
```{r ppc_biv1, echo=FALSE, fig.align='center', fig.height=4, fig.width=4, message=FALSE, warning=FALSE, cache=TRUE}
load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/n2o_mod1.rda")
df_model %>%
  add_predicted_draws(n2o_mod1, 
                      ndraws = 20) %>%
  tidyr::pivot_wider(names_from = .category,
              values_from = .prediction) %>%
  ggplot(aes(x = log(n2o), y = log(n2o_eq))) +
  geom_density_2d(aes(x = logn2o, 
                      y = logn2oeq, 
                      group = .draw),
                  bins = 10,
                  color = "lightblue", 
                  alpha = 0.4) +
  geom_density_2d(color = "black", bins = 10) +
  xlim(1.25, 2.75) +
  ylim(1.25, 2.75) +
  geom_abline(intercept = 0, slope = 1, linetype = "dashed") +
  theme_tidybayes()
```

The same bivariate check is shown below for the re-visit data.
```{r ppc_biv1_test, echo=FALSE, fig.align='center', fig.height=4, fig.width=4, message=FALSE, warning=FALSE, cache=TRUE}
load("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/n2o_mod1.rda")
df_test %>%
  add_predicted_draws(n2o_mod1, 
                      ndraws = 20) %>%
  tidyr::pivot_wider(names_from = .category,
              values_from = .prediction) %>%
  ggplot(aes(x = log(n2o), y = log(n2o_eq))) +
  geom_density_2d(aes(x = logn2o, 
                      y = logn2oeq, 
                      group = .draw),
                  bins = 10,
                  color = "lightblue", 
                  alpha = 0.4) +
  geom_density_2d(color = "black", bins = 10) +
  xlim(1.25, 2.75) +
  ylim(1.25, 2.75) +
  geom_abline(intercept = 0, slope = 1, linetype = "dashed") +
  theme_tidybayes()
```

@@

$$\dfrac{N_2O_{diss}} {N_2O_{equi}}$$

@@

```{r ppc_sat1, echo=FALSE, fig.align='center', fig.height=8, fig.width=8, message=FALSE, warning=FALSE, cache=TRUE}

@@

```{r ppc_sat2, echo=FALSE, fig.align='center', fig.height=8, fig.width=8, message=FALSE, warning=FALSE, cache=TRUE}

@@

```{r ppc_sat_check_mod_n2o3, echo=FALSE, fig.align='center', fig.height=8, fig.width=8, message=FALSE, warning=FALSE, cache=TRUE}

@@

```{r ppc_sat_check_mod_n2o6, echo=FALSE, fig.align='center', fig.height=8, fig.width=8, message=FALSE, warning=FALSE, cache=TRUE}
